\HeaderA{dsgnsum}{Summarize the Sites Selected for a Survey Design}{dsgnsum}
\keyword{survey}{dsgnsum}
\begin{Description}\relax
This function summarizes the sites selected for a survey design by producing
contingency tables containing the cross-tabluation of number of sites for
survey design variables and, optionally, for auxiliary variables.
\end{Description}
\begin{Usage}
\begin{verbatim}
dsgnsum(sp.obj, auxvar=NULL)
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{sp.obj}] the sp package object of class "SpatialPointsDataFrame" produced
by the grts function that contains survey design information and additional
attribute (auxiliary) variables.
\item[\code{auxvar}] a vector containing the names of columns from sites that
identify auxiliary variables to be used to summarize the survey design.
\end{ldescription}
\end{Arguments}
\begin{Value}
A list containing the following components:
\begin{ldescription}
\item[\code{DesignSum}] a list of contingency tables containing the cross-tabulation
of number of sites for the following combinations of survey design
variables:\\
(1) multidensity category (mdcaty) and stratum\\
(2) stratum and panel\\
(3) mdcaty, panel, and stratum
\item[\code{AuxVarSum}] a list of contingency tables containing the cross-tabulation
of number of sites for each auxiliary variable and the design variables
mdcaty, panel, and stratum
\end{ldescription}

In addition the output list plus labeling information is printed to the
console.
\end{Value}
\begin{Author}\relax
Tony Olsen \email{Olsen.Tony@epa.gov}\\
Tom Kincaid \email{Kincaid.Tom@epa.gov}
\end{Author}
\begin{References}\relax
Stevens, D.L., Jr., and A.R. Olsen. (2004). Spatially-balanced sampling of
natural resources. Journal of the American Statistical Association \bold{99}:
262-278.
\end{References}
\begin{SeeAlso}\relax
\code{\LinkA{grts}{grts}}
\code{\LinkA{framesum}{framesum}}
\end{SeeAlso}
\begin{Examples}
\begin{ExampleCode}
## Not run: 
test.design <- list(Stratum1=list(panel=c(PanelOne=50),
   seltype="Equal", over=10), Stratum2=list(panel=c(PanelOne=50,
   PanelTwo=50), seltype="Unequal", caty.n=c(CatyOne=25, CatyTwo=25,
   CatyThree=25, CatyFour=25), over=75)
test.attframe <- read.dbf("test.shapefile")
test.sample <- grts(design=test.design, DesignID="Test.Site", type.frame="area",
   src.frame="shapefile", in.shape="test.shapefile", att.frame=test.attframe,
   stratum="test.stratum", mdcaty="test.mdcaty", shapefile=TRUE,
   shapefilename="test.sample")
dsgnsum(test.sample, auxvar=c("test.ecoregion", "test.state"))
## End(Not run)
\end{ExampleCode}
\end{Examples}

