\HeaderA{write.object}{Write an Object to a Plot}{write.object}
\keyword{aplot}{write.object}
\begin{Description}\relax
This function writes the contents of an object, which may be either a data frame
or a matrix, to a plot.
\end{Description}
\begin{Usage}
\begin{verbatim}
write.object(obj, n.digits=2, r.names=TRUE, c.names=TRUE, r.cex=1, c.cex=1,
   miss="NA")
\end{verbatim}
\end{Usage}
\begin{Arguments}
\begin{ldescription}
\item[\code{obj}] the object (either a data frame or a matrix).
\item[\code{n.digits}] the number of digits after the decimal point for numeric
values.  The default is 2.
\item[\code{r.names}] a logical value that indicates whether to print the row names,
where TRUE = print the row names and  FALSE = do not print the row names.
The default is TRUE.
\item[\code{c.names}] a logical value that indicates whether to print the column
names, where TRUE = print the column names and  FALSE = do not print the
column names.  The default is TRUE.
\item[\code{r.cex}] character expansion parameter for the row labels.  The default is
1.
\item[\code{c.cex}] character expansion parameter for the column labels.  The default
is 1.
\item[\code{miss}] the missing value code expressed as a character string.  The
default is "NA".
\end{ldescription}
\end{Arguments}
\begin{Value}
NULL
\end{Value}
\begin{Author}\relax
Tom Kincaid \email{Kincaid.Tom@epa.gov}
\end{Author}
\begin{Examples}
\begin{ExampleCode}
z <- rnorm(100)
z.mean <- c(tapply(z, rep(1:4, rep(25,4)), mean), mean(z))
z.sd <- sqrt(c(tapply(z, rep(1:4, rep(25,4)), var), var(z)))
z.upper <- z.mean+1.96*z.sd
z.lower <- z.mean-1.96*z.sd
obj <- data.frame(rbind(z.mean, z.sd, z.upper, z.lower))
dimnames(obj) <- list(c("Mean Estimate", "Standard Deviation",
  "Lower 95% Conf. Bound", "Upper 95% Conf. Bound"), c(
  paste("Stratum", 1:4, sep=""), "AllStrata"))
write.object(obj, n.digits=3, r.cex=0.75)

obj <- data.frame(matrix(round(5 + runif(30), 1), nrow=6))
colnames(obj) <- c("United.States", "Russia", "Germany",
  "Japan", "France")
write.object(obj, n.digits=1, r.names=FALSE)
\end{ExampleCode}
\end{Examples}

